---
title: "Bayesian vs Frequentist Accumulation Curve CIs"
date: "April 21, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(coda)
library(doParallel)
library(ggplot2)
load("empirical_probs.RDATA")
source("convenience_functions.r")
```

## Accumulation Curve CI

### Frequentist confidence interval

```{r}
int.list.burd <- list(length = ncol(probs.Burd) - 1)
for(i in 2:ncol(probs.Burd)) {
  probs <- probs.Burd[, i] 
  hit.vec <- probs.Burd$Observed 
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx] 
  
  m <- length(probs)
  
  # Matrix containing the number of hits and lower and upper bounds for 95%
  # confidence intervals for each number of tests
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  for(j in 1:m) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    int.mat[j, 2:3] <- j * CPInt(x = sum(hit.vec[1:j]), p.vec = probs[1:j])
  }
  int.list.burd[[i-1]] <- int.mat
}
```

### Bayesian credible intervals with Jeffreys prior

I am using the model that I described today.  Let me know if you would like
more details.

Blue CIs are Bayesian, Red are Frequentist.

```{r, cache=T}

modelType <- c("Tree", "RF", "SVM", "NNet", "KNN", "PLSLDA")

u <- 1:500
i <- 2
for (i in 2:7) {
  
  hit.vec <- probs.Burd$Observed 
  probs <- probs.Burd[, i]
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx]
  
  m <- length(probs)
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  a <- vector(length = 500)
  b <- vector(length = 500)
  for(j in 1:500) {
    a[j] <- sum(hit.vec[1:j]) + .5
    b[j] <- j - sum(hit.vec[1:j]) + .5
  }
  
  sum.samp <- vector(length = 100000)
  for(j in 1:500) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    sum.samp <- rbeta(100000, a[j], b[j]) * j
    int.mat[j, 2] <- quantile(sum.samp, probs = .025)
    int.mat[j, 3] <- quantile(sum.samp, probs = .975)
  }

  plot(int.list.burd[[i-1]][, 1], type = "l", ylim = c(0, 75),
       main = colnames(probs.Burd)[i+1], ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.list.burd[[i-1]][, 2], type = "l", lty = "dashed", col = "red")
  lines(int.list.burd[[i-1]][, 3], type = "l", lty = "dashed", col = "red")
  
  lines(int.mat[, 2], type = "l", lty = "dashed", col = "blue")
  lines(int.mat[, 3], type = "l", lty = "dashed", col = "blue")
}


```

### Bayesian credible intervals with more informative prior

Early on in the curve I am now putting much higher probability that $\theta_{ij}$ is
close to 1.  Later on in the curve the sample size is so large that the prior doesn't
have any influence on the posterior, so I am using close to the Jeffrey's prior at this 
point.

Blue CIs are Bayesian, Red are Frequentist.

```{r, cache=T}

modelType <- c("Tree", "RF", "SVM", "NNet", "KNN", "PLSLDA")

u <- 1:500
i <- 2
for (i in 2:7) {
  
  hit.vec <- probs.Burd$Observed 
  probs <- probs.Burd[, i]
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx]
  
  m <- length(probs)
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  a <- vector(length = 500)
  b <- vector(length = 500)
  for(j in 1:500) {
    a[j] <- sum(hit.vec[1:j]) + .5 + .005 * (500-i)
    b[j] <- j - sum(hit.vec[1:j]) + .5
  }
  
  sum.samp <- vector(length = 10000)
  for(j in 1:500) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    sum.samp <- rbeta(10000, a[j], b[j]) * j
    int.mat[j, 2] <- quantile(sum.samp, probs = .025)
    int.mat[j, 3] <- quantile(sum.samp, probs = .975)
  }

  plot(int.list.burd[[i-1]][, 1], type = "l", ylim = c(0, 75),
       main = colnames(probs.Burd)[i+1], ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.list.burd[[i-1]][, 2], type = "l", lty = "dashed", col = "red")
  lines(int.list.burd[[i-1]][, 3], type = "l", lty = "dashed", col = "red")
  
  lines(int.mat[, 2], type = "l", lty = "dashed", col = "blue")
  lines(int.mat[, 3], type = "l", lty = "dashed", col = "blue")

}

```


