---
title: "Bayesian vs Frequentist Accumulation Curve CIs"
date: "April 21, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(coda)
library(doParallel)
library(ggplot2)
library(formattable)
library(kableExtra)
library(knitr)
library(magick)
load("empirical_probs.RDATA")
source("convenience_functions.r")
```

## Accumulation Curve CI

The credibility intervals for different machine-learning models and Burden/Pharmacophore descripters. First with the prior being dependent on the ordered indexing of the compounds selected: $Beta( .01, 0.01 + .0005*u)$, where $i$ is the number of compounds selected.

For the Burden Descriptors
```{r, cache=T}

modelType <- c("Tree", "RF", "SVM", "NNet", "KNN", "PLSLDA")

u <- 1:500

int.list.burd.Bayes <- list(length = ncol(probs.Burd) - 1)

for (i in 2:7) {
  
  hit.vec <- probs.Burd$Observed 
  probs <- probs.Burd[, i]
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx]
  
  m <- length(probs)
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  a <- hit.vec + .01
  b <- 1 - hit.vec + .01 + .0005*u
  
  sum.samp <- vector(length = 10000)
  for(j in 1:500) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    sum.samp <- sum.samp + rbeta(10000, a[j], b[j])
    int.mat[j, 2] <- quantile(sum.samp, probs = .025)
    int.mat[j, 3] <- quantile(sum.samp, probs = .975)
  }
  
  par(mfrow = c(1, 1))
  plot(int.mat[, 1], type = "l", ylim = c(0, 100),
       main = paste("Burden: ", colnames(probs.Burd)[i]), ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.mat[, 2], type = "l", lty = "dashed", col = "red")
  lines(int.mat[, 3], type = "l", lty = "dashed", col = "red")
  
  int.list.burd.Bayes[[i-1]] <- int.mat

}
```

For the Pharmocophores
```{r, cache=T}

int.list.phar.Bayes <- list(length = ncol(probs.Phar) - 1)

for (i in 2:7) {
  
  hit.vec <- probs.Phar$Observed 
  probs <- probs.Phar[, i]
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx]
  
  m <- length(probs)
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  a <- hit.vec + .01
  b <- 1 - hit.vec + .01 + .0005*u
  
  sum.samp <- vector(length = 10000)
  for(j in 1:500) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    sum.samp <- sum.samp + rbeta(10000, a[j], b[j])
    int.mat[j, 2] <- quantile(sum.samp, probs = .025)
    int.mat[j, 3] <- quantile(sum.samp, probs = .975)
  }
  
  par(mfrow = c(1, 1))
  plot(int.mat[, 1], type = "l", ylim = c(0, 100),
       main = paste("Pharmacophores: ", colnames(probs.Phar)[i]), ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.mat[, 2], type = "l", lty = "dashed", col = "red")
  lines(int.mat[, 3], type = "l", lty = "dashed", col = "red")
  
  int.list.phar.Bayes[[i-1]] <- int.mat
  
}
```

And then the frequentist runs

Burden
```{r, cache=T}
int.list.burd.Freq <- list(length = ncol(probs.Burd) - 1)
for(i in 2:ncol(probs.Burd)) {
  probs <- probs.Burd[, i] 
  hit.vec <- probs.Burd$Observed 
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx] 
  
  m <- length(probs)
  
  # Matrix containing the number of hits and lower and upper bounds for 95%
  # confidence intervals for each number of tests
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  
  for(j in 1:m) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    int.mat[j, 2:3] <- j * CPInt(x = sum(hit.vec[1:j]), p.vec = probs[1:j])
  }
  int.list.burd.Freq[[i-1]] <- int.mat
}


# Plotting the accumulation curves and confidence band for each modeling method
par(mfrow = c(2, 3))
for (i in seq_along(int.list.burd.Freq)) {
  plot(int.list.burd.Freq[[i]][, 1], type = "l", ylim = c(0, 75),
       main = colnames(probs.Burd)[i+1], ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.list.burd.Freq[[i]][, 2], type = "l", lty = "dashed", col = "red")
  lines(int.list.burd.Freq[[i]][, 3], type = "l", lty = "dashed", col = "red")
}
```

Pharmacophore
```{r, cache=T}
# Confidence intervals for the accumulation curves for Pharmamocophores descriptors Models

int.list.phar.Freq <- list(length = ncol(probs.Burd) - 1)
for(i in 2:ncol(probs.Phar)) {
  probs <- probs.Phar[, i]
  hit.vec <- probs.Phar$Observed
  
  order.idx <- order(probs, decreasing = T)
  probs <- probs[order.idx]
  hit.vec <- hit.vec[order.idx]

  m <- length(probs)
  int.mat <- matrix(ncol = 3, nrow = m)
  colnames(int.mat) <- c("NHits", "LB", "UB")
  for(j in 1:m) {
    int.mat[j, 1] <- sum(hit.vec[1:j])
    int.mat[j, 2:3] <- j * CPInt(x = sum(hit.vec[1:j]), p.vec = probs[1:j])
  }
  int.list.phar.Freq[[i-1]] <- int.mat
}

par(mfrow = c(2, 3))
for (i in seq_along(int.list.phar.Freq)) {
  plot(int.list.phar.Freq[[i]][, 1], type = "l", ylim = c(0, 75),
       main = colnames(probs.Burd)[i+1], ylab = "Number of Hits",
       xlab = "Number of Compounds Selected")
  lines(int.list.phar.Freq[[i]][, 2], type = "l", lty = "dashed", col = "red")
  lines(int.list.phar.Freq[[i]][, 3], type = "l", lty = "dashed", col = "red")
}
```


```{r}

intervals <- list(Burd.Freq = int.list.burd.Freq, Phar.Freq = int.list.phar.Freq,
                  Burd.Bayes = int.list.burd.Bayes, Phar.Bayes = int.list.phar.Bayes)


sig.difs <- vector(length = 4)
names(sig.difs) <- c("BF", "PF", "BB", "PB")

av.width <- vector(length = 4)
names(av.width) <- c("BF", "PF", "BB", "PB")

av.dist.5 <- vector(length = 4)
names(av.width) <- c("BF", "PF", "BB", "PB")

for (l in 1:4) {
  overlap.count <- 0
  total.count <- 0
  width <- 0
  dist.5 <- 0
  for (i in 1:6) {
    for (k in 1:500) {
      for (j in i:6) {
        if ((intervals[[l]][[i]][k, 3] > intervals[[l]][[j]][k, 2] &
            intervals[[l]][[i]][k, 2] < intervals[[l]][[j]][k, 2]) | 
            (intervals[[l]][[j]][k, 3] > intervals[[l]][[i]][k, 2] &
            intervals[[l]][[j]][k, 2] < intervals[[l]][[i]][k, 2])) {
          overlap.count <- overlap.count + 1
        } 
        total.count <- total.count + 1
      }
      width <- width + intervals[[l]][[i]][k, 3] - intervals[[l]][[i]][k, 2]
      dist.5 <- dist.5 + abs((intervals[[l]][[i]][k, 3] + intervals[[l]][[i]][k, 2])/(k*2) - .5)
    }
  }
  sig.difs[l] <- 1 - overlap.count/total.count
  av.width[l] <- width/(6*500)
  av.dist.5[l] <- dist.5/(6*500)
}
sig.difs
av.width
av.dist.5

df <- data.frame(rbind(sig.difs, av.width, av.dist.5))
df <- round(df, 2)
df <- df[, c(1,3,2,4)]
rownames(df) <- c("Fraction of Signficant Differences", "Average Interval Width", "Average Distance from .5")
colnames(df) <- c("Frequentist", "Bayes", "Frequentist", "Bayes")

kable(df, format = "latex", align = "c", booktabs = T, caption = "Summary measures of all point wise confidence/credible intervals", escape = T) %>% kable_styling(latex_options = "striped", position = "center") %>% add_header_above(c(" ", "Burden Numbers" = 2, "Pharmacophores" = 2), italic = T, bold = T) %>% column_spec(1, bold = T)

sig.difs.freq <- (sig.difs[1] + sig.difs[2])/2
sig.difs.bayes <- (sig.difs[3] + sig.difs[4])/2


```



